---
title: "Advanced Data Visualisation"
author: "Billy Lozowski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages and data}
library(tidyverse)
library(ggpubr)
library(ggrepel)

mycotoxin <-read.csv("~/Academics/2. Auburn/Classes/10. Spring 2025/PLPA 5820 - Principles and Techniques of Reproducible Science/In-Class Practice/InClass_practice/MycotoxinData.csv", na.strings = "na")
str(mycotoxin)

mycotoxin$Cultivar <- as.factor(mycotoxin$Cultivar) # convert Cultivar to factor levels

```


# Question 1

Using ggplot, create a boxplot of DON by Treatment so that the plot looks like the
image below.
  a. Jitter points over the boxplot and fill the points and boxplots Cultivar with two
colors from the cbbPallete introduced last week.
  b. Change the transparency of the
jittered points to 0.6.
  c. The y-axis should be labeled "DON(ppm)", and the x-axis should be
left blank.
  d. The plot should use a classic theme
  e. The plot should also be faceted by Cultivar

```{r Q1}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p1 <- ggplot(mycotoxin, aes(Treatment, DON, fill = Cultivar)) +       # set plot mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) + # dodge the boxplots + hide outliers
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                 # set point shape, dodge width, and transparency
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +   # specify fill colours
  labs(x = "",                                                        # define labels
       y = "DON (ppm)") +
  theme_classic() +
  facet_wrap(~Cultivar)                                               # separate plots by Cultivar
p1

```

# Question 2

Change the factor order level so that the treatment “NTC” is first, followed by “Fg”, “Fg + 37”, “Fg + 40”, and “Fg + 70

```{r Q2}
# set factor levels
mycotoxin$Treatment <- factor(mycotoxin$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

p2 <- ggplot(mycotoxin, aes(Treatment, DON, fill = Cultivar)) +       # set plot mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) + # dodge the boxplots + hide outliers
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                 # set point shape, dodge width, and transparency
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +   # specify fill colours
  labs(x = "",                                                        # define labels
       y = "DON (ppm)") +
  theme_classic() +
  facet_wrap(~Cultivar)                                               # separate plots by Cultivar
p2

```

# Question 3

Change the y-variable to plot X15ADON and MassperSeed_mg. The y-axis label should now be “15ADON” and “Seed Mass (mg)”. Save plots made in questions 1 and 3 into three separate R objects.

```{r Q3}
p3 <- ggplot(mycotoxin, aes(Treatment, X15ADON, fill = Cultivar)) +          # mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +        # boxplot mapping
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                        # jitter mapping
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +          # fill colours
  labs(x = "",                                                               # labels
       y = "15ADON") +
  theme_classic() +
  facet_wrap(~Cultivar)                                                      # separate plots
p3

p4 <- ggplot(mycotoxin, aes(Treatment, MassperSeed_mg, fill = Cultivar)) +   # mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +        # boxplot mapping
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                        # jitter mapping
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +          # fill colours
  labs(x = "",                                                               # labels
       y = "Seed Mass (mg)") +
  theme_classic() +
  facet_wrap(~Cultivar)                                                      # separate plots
p4

```

# Question 4

Use ggarrange function to combine all three figures into one with three columns and one row. Set the labels for the subplots as A, B and C. Set the option common.legend = T within ggarage function. What did the common.legend option do?

__the common.legend option shows Cultivar__

```{r Q4}
fig1 <- ggarrange(          
  p1, p3, p4,             # select plot objects
  labels = "auto",        
  nrow = 1,               # set rows
  ncol = 3,               # set columns
  common.legend = T
  )
fig1

```

# Question 5

Use geom_pwc() to add t.test pairwise comparisons to the three plots made above. Save each plot as a new R object, and combine them again with ggarange as you did in question 4.

```{r Q5}

P1 <- p1 + 
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")

P3 <- p3 +
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")

P4 <- p4 +
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")

# arrange plots + stats in combined figure
fig2 <- ggarrange(
  P1, P3, P4,        # select plot objects
  labels = "auto",
  nrow = 1,
  ncol = 3,
  legend = FALSE     # remove legend
)
fig2

```


