---
title: "Intro to Data Visualisation"
author: "Billy Lozowski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

1a. The three elements required to produce a ggplot are the data, the mapping/aesthetics (x & y), and the geometry (geom_point() / geom_boxplot() etc.).

1b. A geom is the plot type. geom_point() for example produces a scatter plot.

1c. A facet is a panel which separates the data by a specific grouping (i.e. treatment type, gender, country etc.).

1d. Layering is the process of adding different elements to a ggplot. Layers added later are added on top ofearlier ones, which need to be considered if a specfic aspect of the plot is more important than another (i.e. a lm line over scatter data).

1e. x and y variables, as well as colour/shape arguments are placed within the aes() aspect of the plot.

## Question 2

```{r import data}
library(tidyverse)
file_path <- file.choose()
df_myo <- read.csv(file_path, na.strings = "na")
glimpse(df_myo)

```

```{r boxplots}
ggplot(df_myo, aes(Treatment, DON, colour = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  theme_classic()

```

## Question 3

```{r bar charts}
ggplot(df_myo, aes(Treatment, DON, colour = Cultivar)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
  xlab("") +
  ylab("DON (ppm)") +
  theme_classic()

```

## Question 4

```{r boxplot and scatter points}
ggplot(df_myo, aes(Treatment, DON, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  geom_point(shape = 21, position = position_jitterdodge()) +
  theme_classic()

```

## Question 5

```{r}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(df_myo, aes(Treatment, DON, color = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  geom_point(shape = 21, position = position_jitterdodge()) +
  scale_color_manual(values = cbbPalette) +
  theme_classic()

```

## Question 6

```{r faceted}
ggplot(df_myo, aes(Treatment, DON, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  geom_point(shape = 21, position = position_jitterdodge()) +
  facet_wrap(~Cultivar) +
  theme_classic()

```

## Question 7

```{r adjust the alpha}
ggplot(df_myo, aes(Treatment, DON, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab("DON (ppm)") +
  geom_point(shape = 21, position = position_jitterdodge(), alpha = 0.5) +
  facet_wrap(~Cultivar) +
  theme_classic()

```

# Question 8

```{r ridge plot}
# install.packages("ggridges")
library(ggridges)
ggplot(df_myo, aes(x = DON, y = Treatment, fill = Cultivar)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none") +
  xlab("DON (ppm)") +
  ylab("") + 
  facet_wrap(~Cultivar, scales = "free")
```






